Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy

### 2023-11-26 Text/Articles/MIGARS_2024/R1-The Entropy as a Proxy for Fully Developed Speckle.Rmd
Rename and create new working version

One of the earliest non-parametric estimators relying on spacings was introduced by Vasicek&nbsp;[@vasicek1976test]. Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* Correa [@correa1995new]: $\widehat{H}_{\text{C}}$.

* Ebrahimi [@Ebrahimi1994]: $\widehat{H}_{\mathrm{E}}(\bm{Z})$.

* Al-Omari [@IbrahimAlOmari2014]: $\widehat{H}_{\mathrm{AO}}(\bm{Z})$.

These estimators, along with others, are described and studied in&nbsp;[@Cassetti2022].

@@

Let's assume that non-parametric entropy estimators \(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, that is:
\begin{equation}
\label{Eq:bias1}
\bias\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta.
\end{equation}
Our objective is to devise unbiased estimators with reduced variance. To achieve this, we introduce an "ideal estimator" \(\check{\theta}(\bm{Z})\) using the bias information:
\begin{equation}
\label{Eq:bias2}
\widecheck{\theta}(\bm{Z}) = \widehat{\theta}(\bm{Z}) - \bias\big(\widehat{\theta}(\bm{Z})\big).
\end{equation}
However, \(\check{\theta}(\bm{Z})\) is not an estimator, because it does not depend on the true parameter \(\theta\), prompting the formulation of a new unbiased  estimator \(\widetilde{H}\), from \eqref{Eq:bias1} and \eqref{Eq:bias2} we have:
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*}
where $B$ represents the number of replications in the bootstrap technique.
Applying this methodology, the original estimators by Correa, Ebrahimi, and Al-Omari are now denoted as the proposed bootstrap-enhanced versions: $\widetilde{H}_{\text{C}}$, $\widetilde{H}_{\text{E}}$, and $\widetilde{H}_{\text{AO}}$, respectively.

@@

Simulations are conducted using $\mathcal{G}_I^0$ distribution, with 200 synthetic samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\), with parameters \(\mu \in \left\{1, 10\right\}\), \(\alpha=-20\), and \(L=8\). In the case of the bootstrap technique, each sample is replicated 100 times through bootstrapping.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $\\mathcal{G}_I^0$, $L=8$, $\\alpha=-20$.", fig.width=6,fig.height=5}
# Define sample sizes for analysis
set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

@@

A hypothesis testing was conducted using non-parametric entropy estimators as test statistics. The results in &nbsp;\ref{tab:table_hipotesis} show that data from the $\mathcal{G}_I^0$ distribution exhibit fully developed speckle behavior, specifically in the limit case with parameters $\mu=10$, $L=5$, and $\alpha= -400$. The true entropy of $\Gamma_{\text{SAR}}$ was set at $H_{\Gamma_{\text{SAR}}} = 2.846$.  It is observed that, for different sample sizes, entropy values converge towards the true value of the Gamma SAR distribution. Hypothesis test results were conducted with a 95% confidence level. The $Z$ statistic measures the discrepancy between the estimated entropy and the true entropy. The $p$-values associated with the $Z$ statistic are predominantly greater than $0.05$, for sample sizes above 25, suggesting that the data is consistent with the null hypothesis.

>>>>

One of the earliest non-parametric estimators relying on spacings was introduced by @vasicek1976test. 
Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* @correa1995new: $\widehat{H}_{\text{C}}$.

* @Ebrahimi1994: $\widehat{H}_{\mathrm{E}}(\bm{Z})$.

* @IbrahimAlOmari2014: $\widehat{H}_{\mathrm{AO}}(\bm{Z})$.

These estimators, along with others, are described and studied in @Cassetti2022.

@@

Let's assume that non-parametric entropy estimators \(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, that is:
\begin{equation}
\label{Eq:bias1}
\bias\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta.
\end{equation}
Our objective is to devise unbiased estimators with reduced variance. To achieve this, we introduce an "ideal estimator" \(\check{\theta}(\bm{Z})\) using the bias information:
\begin{equation}
\label{Eq:bias2}
\widecheck{\theta}(\bm{Z}) = \widehat{\theta}(\bm{Z}) - \bias\big(\widehat{\theta}(\bm{Z})\big).
\end{equation}
However, \(\check{\theta}(\bm{Z})\) is not an estimator, because it depends on the true parameter \(\theta\), prompting the formulation of a new estimator \(\widetilde{H}\). 
From \eqref{Eq:bias1} and \eqref{Eq:bias2} we have:
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*}
where $B$ is the number of replications in the bootstrap technique.
Applying this methodology, the original estimators by Correa, Ebrahimi, and Al-Omari are now denoted as the proposed bootstrap-enhanced versions: $\widetilde{H}_{\text{C}}$, $\widetilde{H}_{\text{E}}$, and $\widetilde{H}_{\text{AO}}$, respectively.

@@

Simulations are conducted using $\mathcal{G}_I^0$ distribution, with $200$ simulated samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\), with parameters \(\mu \in \left\{1, 10\right\}\), \(\alpha=-20\), and \(L=8\). In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $\\mathcal{G}_I^0$, $L=8$, $\\alpha=-20$.", fig.width=6, fig.height=5}
# Define sample sizes for analysis
set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

@@

A hypothesis testing was conducted using non-parametric entropy estimators as test statistics. 
The results in Table&nbsp;\ref{tab:table_hipotesis} show that data from the $\mathcal{G}_I^0$ distribution exhibit fully developed speckle behavior, specifically in the limit case with parameters $\mu=10$, $L=5$, and $\alpha= -400$. 
The true entropy of $\Gamma_{\text{SAR}}$ was set at $H_{\Gamma_{\text{SAR}}} = 2.846$.  
It is observed that, for different sample sizes, entropy values converge towards the true value of the Gamma SAR distribution. 
Hypothesis test results were conducted with a \SI{95}{\percent} confidence level. 
The $Z$ statistic measures the discrepancy between the estimated entropy and the true entropy. 
The $p$-values associated with the $Z$ statistic are predominantly greater than $0.05$, for sample sizes above $25$, suggesting that the data are consistent with the null hypothesis.
